Parsing Tweets into Universal Dependencies

### 2017-12-15 paper.tex
fix and update.

Kong et al.'s rapid, small-scale annotation effort was heavily constrained.  It was
carried out by annotators with only cursory training, no clear
annotation guidelines, and no effort to achieve consensus on controversial
cases. Annotators were allowed to underspecify their analyses.
Most of the work was done in a very short amount of
time (a day).  Driven both by the style of the text they sought to annotate
and by exigency, some of their annotation conventions included:
\begin{itemize}
\item Allowing an annotator to exclude tokens from the dependency
  tree.  A clear criterion for exclusion was not given, but many
  tokens were excluded because they were deemed ``non-syntactic.'' \yicomment{including some Twitter specific tokens, which we will refine the definition in Section 2.3 and 2.4?}
\item Allowing an annotator to merge a multiword expression into a
  single node in the dependency tree, with no internal structure.
  Annotators were allowed to take the same step with noun phrases.
\item Allowing multiple roots, since a single tweet might contain more
  than one sentence.
\end{itemize}
These conventions were justified on the grounds of making the
annotation easier for non-experts, but they must be revisited in our
effort to apply UD to tweets.

@@

The major characteristic that distinguishes tweets from standard texts is that
a large proportion of tokens don't carry any syntactic function. 
Apart from conventional non-syntactic tokens appearing in both types of texts, such as punctuations and exclamations,
in our annotation, there are still five types of non-syntactic tokens tweets have\yicomment{need to say it explicitly that here we only refer to Twitter/web specific tokens, not including exclamations or punctuations or other stuffs. Or we can say punctuations and exclamations are canonically non-syntacitc}: 
sentiment emoticons, retweet markers and their following at-mentions, topical hashtags, referential URLs, and
truncated words.\footnote{The tweets we analyze have at most 140
  characters. Although Twitter has doubled the tweet length limit to
  280 characters since our analysis, we believe this type of token
  will still remain.}  Figure \ref{fig:non-syn-toks} illustrates examples of
these non-syntactic tokens.  As discussed above, these are generally
tagged with the \emph{other} (\texttt{X}) part of speech, except
emoticons, which are tagged as \emph{symbol} (\texttt{SYM}).  In our annotation, 
7.55\% of all tokens are belong to one of the five types; detailed statistics can be found in Table \ref{tbl:non-synt-prop}.
%\yjcomment{I've merged retweet marker and retweet at-mention into one.}

@@

Inspired by the way UD deals with \emph{punctuation} (which is
canonically non-syntactic), we adopt the following
conventions\yicomment{this applies to all non-syntactic tokens including punctuations and excalamations}:
%\yicomment{shall we mention how we segment sentences? Seems relevant
%here}\yjcomment{i don't want to touch sentence segmentation, it will
%makes the paper more messy.}\nascomment{agree with Yijia}

\begin{itemize}
%\item For the at-mentions in retweet constructions, we attach them to their retweet markers. We will discuss our choice in the following paragraph;
\item If a non-syntactic token is within a sentence that has a clear predicate, it will be attached to this predicate. Retweet construction is a special case and we will discuss our choice in the following paragraph;
\item If the whole sentence is made of a sequence of non-syntactic tokens, we attach all these tokens to the first one;
%\item In parataxis case, when it is difficult to decide which sentence or phrase x should belong to, we always attach x to the previous predicate of the paratactic clause or phrase;
\item Non-syntactic tokens are mostly labeled as {\it discourse}, but
  URLs are always labeled as {\it list}, following the UD\_English dataset. \yicomment{need to add punct for punctuations?}
%\\ \yicomment{For retweet at-mention case, we attach the retweet at-mentions to the RT sign, so that the retweet action and the people whose tweet is being retweeted could be connected.\footnote{We can also follow the first convention to attach the retweet at-mention to its predicate, but we think it is better for downstream tasks.}}
\end{itemize}

@@

\paragraph{Retweet construction.} Figure \ref{fig:non-syn-toks} shows
an example of the retweet construction (\emph{RT @coldplay :}).  This
might be treated as a verb phrase, with \emph{RT} as a verb and the
at-mention as its object\yicomment{@coldplay seems not to be an object, could be some modifier like vocative, or some prepositional case (the tweets on/about coldplay)}.
% \yicomment{or might be treated as 2 non-syntactic tokens and a punctuation, which should all attach to the predicate of the following retweeted tweet.}
%\yjcomment{i wouldn't like the readers to think in this direction.}.
  This solution would lead to an uninformative root
word and, since this expression is idiomatic to Twitter, might create
unnecessary confusion for downstream applications aiming to identify
the main predicate(s) of a tweet.  We therefore treat the whole
expression as non-syntactic, including assigning the \emph{other}
(\texttt{X}) part of speech to both \emph{RT} and \emph{@coldplay},
attaching the at-mention to \emph{RT} with the \emph{discourse}
label and the colon to \emph{RT} with the \emph{punct}(uation) label,
and attaching \emph{RT} to the predicate of the following sentence.
%\yicomment{since we have introduced retweet at-mentions in the non-syntactic tokens, here we can just briefly justify our choice.}.

@@

\paragraph{Vocative at-mentions.}  Another idiomatic construction on
Twitter is a vocative at-mention (sometimes a
signal that a tweet is a reply to a tweet by the mentioned user). % \yicomment{not necessary, can be just mentioning somebody. should be careful to say it. We can say this construction include 1. reply;  2. mention sb. Both cases are similar to general vocative proper noun case in standard texts. And we need to exclude at-mentions in Retweet Construction.}\yjcomment{i didn't see any difference between reply and mention, they all starts with a @username.}.
We
treat these at-mentions as vocative expressions, labeling them with
POS tag
\emph{proper noun} (\texttt{PROPN}) and attaching them to the
following predicate\yicomment{not necessarily following, better to be `the predicate of the sentence it belongs to'} with the label \emph{vocative} as in UD guidelines (see Figure \ref{fig:informal-toks} for an example).

@@

\subsubsection{Annotation Process}
\begin{table}
	\centering
	\begin{tabular}{rrrr}
		\hline
		& train & dev & test \\
		\hline
		tweets 		&	1,639 		&	709		&	1,201\\
		& +1,000 & +709 & +1,000\\
		tokens 		&	24,753		&	11,742	&	19,112\\
		\hline
	\end{tabular}
\caption{Statistics of {\sc Tweebank v2}.
The second row shows the number of new tweets compared to
{\sc Tweebank v1}.\label{tbl:data-stat}
}
\end{table}
Our annotation process was conducted in two stages.
In the first stage, 18 researchers worked on the {\sc Tweebank v1}
proportion\yicomment{and the additional 210 tweets} and created the initial annotations in one day.
Before annotating, they were given a tutorial overview of the general UD
annotation conventions and our guidelines specifically for annotating tweets.
Both the guidelines and annotations
were further refined by the authors of this paper to increase
the coverage of our guidelines and solve inconsistencies between
different annotators during this exercise. In the second stage, a tokenizer, a POS tagger, and a
parser were trained on the annotated data from the first stage (1,050 tweets in total),
and used to automatically analyze the sampled 2,500 tweets.  Authors 
of this paper manually corrected the parsed data and finally achieved 3,550 labeled tweets.
Newly created annotations are splitted into train, development, and test set and appended
to the original split of {\sc Twebank v1}. Statistics of our annotations and data split are shown
in Table \ref{tbl:data-stat}.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it System & $F_1$ \\
		\hline
		  Stanford CoreNLP & 96.6 \\
		 Twokenizer & 94.4 \\		 
% 		UD pipe v1.2 & 94.7 \\
		\hdashline
		UD pipe v1.2 & 97.3 \\
	 	our bi-LSTM tokenizer & 98.3 \\
		\hline
	\end{tabular}
	\caption{Tokenizer comparison on the \textsc{Tweebank v2} test set.}\label{tbl:tok-result}
\end{table}

We trained our tokenizer on the training portion of
\textsc{Tweebank v2} combined with the UD\_English training set
and tested on the \textsc{Tweebank v2} test set. 
We report $F_1$ scores, combining precision and recall for token identification. Table \ref{tbl:tok-result} shows the
tokenization results, compared to  other available tokenizers. 
 Stanford CoreNLP \cite{manning-EtAl:2014:P14-5} and Twokenizer
\cite{ICWSM101540} systems are rule-based systems and were not adapted
to the UD tokenization scheme.
The UD pipe v1.2
\cite{straka-strakova:2017:K17-3} model was re-trained on the same
data as our system. Compared with the UD pipe, we use LSTM
instead of GRU in our model and we also use a larger size of hidden units (64 against 20),
which has stronger representation power. 
Our bi-LSTM tokenizer achieves the best accuracy among all these
tokenizers.  These results indicate the value of statistical modeling
in tokenization for informal texts.

@@

Part-of-speech tagging for tweets has been extensively studied \cite{ritter-EtAl:2011:EMNLP,gimpel-EtAl:2011:ACL-HLT2011,derczynski-EtAl:2013:RANLP-2013,owoputi-EtAl:2013:NAACL-HLT,gui-EtAl:2017:EMNLP20172}. We therefore consider existing POS taggers for tweets instead of developing our own.
On the annotation scheme designed in \S\ref{sec:pos-anno}, based on UD and adapted for
Twitter, we compared several existing systems: the Stanford CoreNLP tagger, 
\citet{owoputi-EtAl:2013:NAACL-HLT}'s word cluster enhanced tagger
(both greedy and CRF variants), and 
\citet{ma-hovy:2016:P16-1}'s neural network tagger 
which achieves the state-of-the-art performance on PTB.
\citet{gui-EtAl:2017:EMNLP20172} presented the state-of-the-art neural tagger for Twitter,
but their code didn't work beyond PTB tag set. Thus, we didn't include their tagger in our comparison.\yjcomment{i put some words on \cite{gui-EtAl:2017:EMNLP20172} in case the reviewer challenge our neural/non-neural comparsion.}
 %and also two bi-LSTM systems similar
%to \citet{DBLP:journals/corr/HuangXY15} and \citet{lample-EtAl:2016:N16-1}, one using
%word vectors and the other using word and character vectors (see Figure \ref{fig:pos-model}).
All compared systems were re-trained on the combination of the UD\_English and 
\textsc{Tweebank v2} training sets\yicomment{should mention that combination of corpora improves the results, otherwise we should answer why we didn't try training ONLY on Tweebank v2}. We use Twitter-specific GloVe embeddings released by
\citet{pennington-socher-manning:2014:EMNLP2014} in all neural taggers
and parsers.\footnote{\url{http://nlp.stanford.edu/data/glove.twitter.27B.zip}}

@@

\citet{reimers-gurevych:2017:EMNLP2017} and others have
pointed out that neural network training is 
nondeterministic and depends on the seed for the random
number generator.
Our preliminary experiments confirm this finding, with a gap of 1 LAS on development data
between the best (75.8)\yicomment{is this for char stacklstm parser? in table 6 it shows the same number. If it's average, then number in tab 6 should $< 75.8$?} and worst (74.8) runs. To control for this
effect, we report the average of five differently-seeded runs, for
each of our models and the compared ones.

\begin{table}[t]
	\centering
	\begin{tabular}{rccc}
		\hline
		\it System & UAS & LAS & Kt/s \\
		\hline
		\citet{kong-EtAl:2014:EMNLP2014} &81.6 & 77.2 & 0.3 \\
		\citet{dozat-qi-manning:2017:K17-3} & 81.9 &  77.7 & 1.7 \\
%		\newcite{dyer-EtAl:2015:ACL-IJCNLP} & 77.9 & 73.0 \\
		\newcite{ballesteros-EtAl:2016:EMNLP2016} & 80.4 & 75.8 & 2.3 \\
		\hdashline
		Ensemble (20) & 83.5 & 79.4 & 0.2 \\
		Distillation ($\alpha =1.0$) & 82.2 & 78.0 & 2.3 \\
		Distillation ($\alpha =0.9$) & 82.4 & 78.1 & 2.3 \\		
		Distillation w/ exploration & 82.5 & 78.4 & 2.3 \\
		\hline
	\end{tabular}
	\caption{Dependency parser comparison on \textsc{Tweebank v2},
           with automatic POS tags. The ``Kt/s'' column shows
           the parsing speed evaluated by thousands of tokens
           the model processed per second. For fair comparison, we limit the
           number of CPUs used in \citet{dozat-qi-manning:2017:K17-3} experiments to 1.
            \label{tbl:parse-result}}
\end{table}

>>>>

Kong et al.'s rapid, small-scale annotation effort was heavily constrained.  It was
carried out by annotators with only cursory training, no clear
annotation guidelines, and no effort to achieve consensus on controversial
cases. Annotators were allowed to underspecify their analyses.
Most of the work was done in a very short amount of
time (a day).  Driven both by the style of the text they sought to annotate
and by exigency, some of their annotation conventions included:
\begin{itemize}
\item Allowing an annotator to exclude tokens from the dependency
  tree.  A clear criterion for exclusion was not given, but many
  tokens were excluded because they were deemed ``non-syntactic.''% \yicomment{including some Twitter specific tokens, which we will refine the definition in Section 2.3 and 2.4?}
\item Allowing an annotator to merge a multiword expression into a
  single node in the dependency tree, with no internal structure.
  Annotators were allowed to take the same step with noun phrases.
\item Allowing multiple roots, since a single tweet might contain more
  than one sentence.
\end{itemize}
These conventions were justified on the grounds of making the
annotation easier for non-experts, but they must be revisited in our
effort to apply UD to tweets.

@@

The major characteristic that distinguishes tweets from standard texts is that
a large proportion of tokens don't carry any syntactic function. 
In our annotation, there are five types of non-syntactic tokens commonly seen in tweets: %\yicomment{need to say it explicitly that here we only refer to Twitter/web specific tokens, not including exclamations or punctuations or other stuffs. Or we can say punctuations and exclamations are canonically non-syntacitc}: 
sentiment emoticons, retweet markers and their following at-mentions, topical hashtags, referential URLs, and
truncated words.\footnote{The tweets we analyze have at most 140
  characters. Although Twitter has doubled the tweet length limit to
  280 characters since our analysis, we believe this type of token
  will still remain.}  Figure \ref{fig:non-syn-toks} illustrates examples of
these non-syntactic tokens.  As discussed above, these are generally
tagged with the \emph{other} (\texttt{X}) part of speech, except
emoticons, which are tagged as \emph{symbol} (\texttt{SYM}).  In our annotation, 
7.55\% of all tokens are belong to one of the five types; detailed statistics can be found in Table \ref{tbl:non-synt-prop}.
%\yjcomment{I've merged retweet marker and retweet at-mention into one.}

@@

Inspired by the way UD deals with \emph{punctuation} (which is
canonically non-syntactic), we adopt the following
conventions%\yicomment{this applies to all non-syntactic tokens including punctuations and excalamations}
:
%\yicomment{shall we mention how we segment sentences? Seems relevant
%here}\yjcomment{i don't want to touch sentence segmentation, it will
%makes the paper more messy.}\nascomment{agree with Yijia}

\begin{itemize}
%\item For the at-mentions in retweet constructions, we attach them to their retweet markers. We will discuss our choice in the following paragraph;
\item If a non-syntactic token is within a sentence that has a clear predicate, it will be attached to this predicate. Retweet construction is a special case and we will discuss our choice in the following paragraph;
\item If the whole sentence is made of a sequence of non-syntactic tokens, we attach all these tokens to the first one;
%\item In parataxis case, when it is difficult to decide which sentence or phrase x should belong to, we always attach x to the previous predicate of the paratactic clause or phrase;
\item Non-syntactic tokens are mostly labeled as {\it discourse}, but
  URLs are always labeled as {\it list}, following the UD\_English dataset.% \yicomment{need to add punct for punctuations?}
%\\ \yicomment{For retweet at-mention case, we attach the retweet at-mentions to the RT sign, so that the retweet action and the people whose tweet is being retweeted could be connected.\footnote{We can also follow the first convention to attach the retweet at-mention to its predicate, but we think it is better for downstream tasks.}}
\end{itemize}

@@

\paragraph{Retweet construction.} Figure \ref{fig:non-syn-toks} shows
an example of the retweet construction (\emph{RT @coldplay :}).  This
might be treated as a verb phrase, with \emph{RT} as a verb and the
at-mention as its object\yicomment{@coldplay seems not to be an object, could be some modifier like vocative, or some prepositional case (the tweets on/about coldplay)}\yjcomment{i am okey with current version}.
% \yicomment{or might be treated as 2 non-syntactic tokens and a punctuation, which should all attach to the predicate of the following retweeted tweet.}
%\yjcomment{i wouldn't like the readers to think in this direction.}.
  This solution would lead to an uninformative root
word and, since this expression is idiomatic to Twitter, might create
unnecessary confusion for downstream applications aiming to identify
the main predicate(s) of a tweet.  We therefore treat the whole
expression as non-syntactic, including assigning the \emph{other}
(\texttt{X}) part of speech to both \emph{RT} and \emph{@coldplay},
attaching the at-mention to \emph{RT} with the \emph{discourse}
label and the colon to \emph{RT} with the \emph{punct}(uation) label,
and attaching \emph{RT} to the predicate of the following sentence.
%\yicomment{since we have introduced retweet at-mentions in the non-syntactic tokens, here we can just briefly justify our choice.}.

@@

\paragraph{Vocative at-mentions.}  Another idiomatic construction on
Twitter is a vocative at-mention (sometimes a
signal that a tweet is a reply to a tweet by the mentioned user). % \yicomment{not necessary, can be just mentioning somebody. should be careful to say it. We can say this construction include 1. reply;  2. mention sb. Both cases are similar to general vocative proper noun case in standard texts. And we need to exclude at-mentions in Retweet Construction.}\yjcomment{i didn't see any difference between reply and mention, they all starts with a @username.}.
We
treat these at-mentions as vocative expressions, labeling them with
POS tag
\emph{proper noun} (\texttt{PROPN}) and attaching them to the
main predicate of the sentence it is within with the label \emph{vocative} as in UD guidelines (see Figure \ref{fig:informal-toks} for an example).\yjcomment{as Yi suggested, change ``following predicate'' to ``main ... within''}

@@

\subsubsection{Annotation Process}
\begin{table}
	\centering
	\begin{tabular}{rrrr}
		\hline
		& train & dev & test \\
		\hline
		tweets 		&	1,639 		&	709		&	1,201\\
		& +1,000 & +709 & +1,000\\
		tokens 		&	24,753		&	11,742	&	19,112\\
		\hline
	\end{tabular}
\caption{Statistics of {\sc Tweebank v2}.
The second row shows the number of new tweets compared to
{\sc Tweebank v1}.\label{tbl:data-stat}
}
\end{table}
Our annotation process was conducted in two stages.
In the first stage, 18 researchers worked on the {\sc Tweebank v1}
proportion and the additional 210 tweets and created the initial annotations in one day.
Before annotating, they were given a tutorial overview of the general UD
annotation conventions and our guidelines specifically for annotating tweets.
Both the guidelines and annotations
were further refined by the authors of this paper to increase
the coverage of our guidelines and solve inconsistencies between
different annotators during this exercise. In the second stage, a tokenizer, a POS tagger, and a
parser were trained on the annotated data from the first stage (1,050 tweets in total),
and used to automatically analyze the sampled 2,500 tweets.  Authors 
of this paper manually corrected the parsed data and finally achieved 3,550 labeled tweets.
Newly created annotations are splitted into train, development, and test set and appended
to the original split of {\sc Twebank v1}. Statistics of our annotations and data split are shown
in Table \ref{tbl:data-stat}.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it System & $F_1$ \\
		\hline
		  Stanford CoreNLP & 96.6 \\
		 Twokenizer & 94.4 \\		 
% 		UD pipe v1.2 & 94.7 \\
		\hdashline
		UD pipe v1.2 & 97.3 \\
	 	our bi-LSTM tokenizer & 98.3 \\
		\hline
	\end{tabular}
	\caption{Tokenizer comparison on the \textsc{Tweebank v2} test set.}\label{tbl:tok-result}
\end{table}
\citet{wang-EtAl:2017:Long6} shows that combining Singlish\footnote{a dialect of English} treebank with UD\_English
can achieve parsing performance on Singlish parsing.
We followed their work and trained our tokenizer on the training portion of
\textsc{Tweebank v2} combined with the UD\_English training set
and tested on the \textsc{Tweebank v2} test set. \yicomment{should mention that combination of corpora improves the results, otherwise we should answer why we didn't try training ONLY on Tweebank v2}
We report $F_1$ scores, combining precision and recall for token identification. Table \ref{tbl:tok-result} shows the
tokenization results, compared to  other available tokenizers. 
 Stanford CoreNLP \cite{manning-EtAl:2014:P14-5} and Twokenizer
\cite{ICWSM101540} systems are rule-based systems and were not adapted
to the UD tokenization scheme.
The UD pipe v1.2
\cite{straka-strakova:2017:K17-3} model was re-trained on the same
data as our system. Compared with the UD pipe, we use LSTM
instead of GRU in our model and we also use a larger size of hidden units (64 against 20),
which has stronger representation power. 
Our bi-LSTM tokenizer achieves the best accuracy among all these
tokenizers.  These results indicate the value of statistical modeling
in tokenization for informal texts.

@@

Part-of-speech tagging for tweets has been extensively studied \cite{ritter-EtAl:2011:EMNLP,gimpel-EtAl:2011:ACL-HLT2011,derczynski-EtAl:2013:RANLP-2013,owoputi-EtAl:2013:NAACL-HLT,gui-EtAl:2017:EMNLP20172}. We therefore consider existing POS taggers for tweets instead of developing our own.
On the annotation scheme designed in \S\ref{sec:pos-anno}, based on UD and adapted for
Twitter, we compared several existing systems: the Stanford CoreNLP tagger, 
\citet{owoputi-EtAl:2013:NAACL-HLT}'s word cluster enhanced tagger
(both greedy and CRF variants), and 
\citet{ma-hovy:2016:P16-1}'s neural network tagger 
which achieves the state-of-the-art performance on PTB.
\citet{gui-EtAl:2017:EMNLP20172} presented the state-of-the-art neural tagger for Twitter,
but their code didn't work beyond PTB tag set. Thus, we didn't include their tagger in our comparison.\yjcomment{i put some words on \cite{gui-EtAl:2017:EMNLP20172} in case the reviewer challenge our neural/non-neural comparsion.}
 %and also two bi-LSTM systems similar
%to \citet{DBLP:journals/corr/HuangXY15} and \citet{lample-EtAl:2016:N16-1}, one using
%word vectors and the other using word and character vectors (see Figure \ref{fig:pos-model}).
All compared systems were re-trained on the combination of the UD\_English and 
\textsc{Tweebank v2} training sets. We use Twitter-specific GloVe embeddings released by
\citet{pennington-socher-manning:2014:EMNLP2014} in all neural taggers
and parsers.\footnote{\url{http://nlp.stanford.edu/data/glove.twitter.27B.zip}}

@@

\citet{reimers-gurevych:2017:EMNLP2017} and others have
pointed out that neural network training is 
nondeterministic and depends on the seed for the random
number generator.
Our preliminary experiments confirm this finding, with a gap of 1 LAS on development data
between the best (75.8)
 and worst (74.8) runs. To control for this
effect, we report the average of five differently-seeded runs, for
each of our models and the compared ones.

\begin{table}[t]
	\centering
	\begin{tabular}{rccc}
		\hline
		\it System & UAS & LAS & Kt/s \\
		\hline
		\citet{kong-EtAl:2014:EMNLP2014} &81.6 & 77.2 & 0.3 \\
		\citet{dozat-qi-manning:2017:K17-3} & 81.9 &  77.7 & 1.7 \\
%		\newcite{dyer-EtAl:2015:ACL-IJCNLP} & 77.9 & 73.0 \\
		\newcite{ballesteros-EtAl:2016:EMNLP2016} & 80.4 & 75.8 & 2.3 \\
		\hdashline
		Ensemble (20) & 83.5 & 79.4 & 0.2 \\
		Distillation ($\alpha =1.0$) & 82.2 & 78.0 & 2.3 \\
		Distillation ($\alpha =0.9$) & 82.4 & 78.1 & 2.3 \\		
		Distillation w/ exploration & 82.5 & 78.4 & 2.3 \\
		\hline
	\end{tabular}
	\caption{Dependency parser comparison on \textsc{Tweebank v2} test set,
           with automatic POS tags. The ``Kt/s'' column shows
           the parsing speed evaluated by thousands of tokens
           the model processed per second. For fair comparison, we limit the
           number of CPUs used in \citet{dozat-qi-manning:2017:K17-3} experiments to 1.
            \label{tbl:parse-result}}
\end{table}
